### 2011-04-13 Text/paper.tex
Writing results. Need to play with VTK.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{tikz}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with rectangular and square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy.
\end{abstract}

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated. To avoid the mutual influence of bubbles on each other the channel length was chosen as $15 H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a body force in the framework of the LBM but not with the pressure difference. Therefore, we can not address simulations with upward and downward flows \citet{cerro-bubble}. It is assumed that the gravity effects are negligible and the flow is in the horizontal direction driven by the body force.  
\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:classical:benchmark}}
\end{figure}[ht]
\begin{figure}
\includegraphics*[bb=152 480 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:lbm:benchmark}}
\end{figure}

@@

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix \ref{appendix}. The next section further examines the steady-state approach, the dependency of the radii on the moderate capillary number values $0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio.

\section{Results}

\section{Steady-state approach}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate.

As well we performed longer simulations in the range of moderate capillary numbers after
$140000$, $160000$, $180000$, $200000$, $220000$, $240000$. The results for the capillary number
based on the bubble velocity in the middle crossection are shown in Table \ref{table:steady:state}. 
%{\color{red} What is it One of the
%profiles in
%the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}. }
\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
$140000$&$0.0532$&$0.9379$&$1.075$\\
$150000$&$0.0533$&$0.9375$&$1.072$\\
$160000$&$0.0535$&$0.9372$&$1.069$\\
$170000$&$0.0536$&$0.9369$&$1.066$\\
$180000$&$0.0537$&$0.9366$&$1.062$\\
\hline
\end{tabular}\\
\begin{tabular}{|c|c|c|c|}
\hline
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
$140000$&$0.3578$&$0.8657$&$0.8708$\\
$160000$&$0.3590$&$0.8650$&$0.8702$\\
$180000$&$0.3598$&$0.8647$&$0.8700$\\
$200000$&$0.3603$&$0.8645$&$0.8698$\\
$220000$&$0.3607$&$0.8644$&$0.8697$\\
$240000$&$0.3609$&$0.8645$&$ 0.8697$\\
\hline
\end{tabular}
\end{center}
\caption{Underresolved capillary number results and resolved Results for the steady-state case. One
can see that $200000$ steps are enough to structure
the bubble. \label{table:steady:state}}
\end{table}

%\begin{figure}
%\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
%\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
%\caption{Slice in the middle of the bubble and
%three-dimensional example. \label{fig:steady:state:profile:example}}
%\end{figure}

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with rectangular and square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy.
\end{abstract}

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:classical:benchmark}}
\end{figure}[ht]
\begin{figure}
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:lbm:benchmark}}
\end{figure}

@@

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
the steady-state approach, the dependency of the radii on the moderate capillary number values
$0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio are
examined, see Section \ref{sec:results}.

\section{Binary liquid lattice Boltzmann model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.

The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and the velocity fields, and another to represent the
phase field $\phi$ indicating gas or liquid.
The equilibrium populations \cite{pooley-contact} are defined as:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, i=1\div Q-1\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div Q-1\\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependent on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
to change viscosity from the gas viscosity
$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
A}{9}}$.

Velocity set is defined as:
\begin{equation} 
\begin{aligned}
&c_{ix}=\{0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1\}\\
&c_{iy}=\{0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0\}\\
&c_{iz}=\{0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1\}.
\end{aligned}
\end{equation}
The weights are $w_0=0$, $w_{1-6}=\frac{1}{6}$ and $w_{7-18}=\frac{1}{12}$. The weights
related to the inclusion of the surface tension are:
\begin{equation}
\begin{aligned}
&w^{xx}_{1-2}=w^{yy}_{3-4}=w^{zz}_{5-6}=\frac{5}{12}\\
&w^{xx}_{3-6}=w^{yy}_{1-2,5-6}=w^{zz}_{1-4}=-\frac{1}{3}\\
&w^{xx}_{7-10,15-18}=w^{yy}_{7-14}=w^{zz}_{11-18}=-\frac{1}{24}\\
&w^{xx}_{11-15,}=w^{yy}_{15-18}=w^{zz}_{7-10}=\frac{1}{12}\\
&w^{xy}_{7,10}=-w^{xy}_{8,9}=w^{yz}_{11,14}=-w^{yz}_{12,13}=w^{zx}_{15,18}=-w^{zx}_{16,17}=\frac{1}{
4}.
\end{aligned}
\end{equation}

\section{Results}
\label{sec:results}
This section describes numerical simulations. First, we examine when the steady-state approach is
achieved. Then, the critical capillary number is identified when the transition from the
non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
number is presented for the range of moderate capillary numbers $0.1 \leq Ca 1.0$. Results are
concluded with studies of the velocity pattern,{\color{red} the aspect ratio, and the influence of
inertia}.

\subsection{Steady-state approach}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with the body force as
$\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
$220000$ and
$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|}
\hline
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
\hline
$140000$&$0.1879$&$0.8657$&$0.8708$\\
$160000$&$0.1878$&$0.8650$&$0.8702$\\
$180000$&$0.1882$&$0.8647$&$0.8700$\\
$200000$&$0.1893$&$0.8645$&$0.8698$\\
$220000$&$0.1925$&$0.8644$&$0.8697$\\
$240000$&$0.1961$&$0.8645$&$0.8697$\\
\hline
\end{tabularx}
\caption{Results for the steady-state case. One
can see that $200000$ steps are enough to structure
the bubble and approach the steady state. The small noise in the capillary number is connected to
the identification of the interface. One needs to address the interpolation and the spurious
currents existing in the system. All other parameters are relaxed to the
steady-state.\label{table:steady:state}}
\end{table}

The bubble velocity is the velocity of the interface. One can see it with ...

\subsection{Radii transition}
Due to performance issues we can only access the
transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
is guaranteed. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
were conducted. Surprisingly the transition occurs in the region computationally accessible for
current simulations. In comparison with the results of the \citet{heil-threedim}
($\widehat{Ca}=0.04$) our
results are more close to the VOF continuous interface method simulation by
\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and experimental dependency of
\citet{shikazono-square}, see Table
\ref{table:transition:results}. The transition happens for $Ca\approx 0.09$.  One can see
two examples of non-symmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see Fig.
\ref{fig:crosssections:sym}.  
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&$R_{axis}$&$R_{diag}$&$Re$&$R_{axis,Wang}$&$R_{diag,Wang}$\\
\hline
$0.053$&$0.95$&$1.01$&$0.907$&$0.99$&$0.99$\\
$0.078$&$0.93$&$0.95$&$1.062$&$0.96$&$0.96$\\
$0.132$&$0.89$&$0.89$&$1.80$&$0.93$&$0.93$\\
\hline
\end{tabularx}
\caption{Simulation results in terms of $R_{axis}$ and $R_{diag}$ for the transition region between
non-symmetric and axisymmetric cases. $R_{axis,Wang}$ and $R_{diag,Wang}$ are experimental
correlations by \citet{wang-non-circular}. \label{table:transition:results}}
\end{table}
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca5.eps}
\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca13.eps}\\
\caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
(left) and $Ca=1.13$ (right). Crosssections are rescaled on the half channel height. Other
parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
is asymmetric in comparison
with the right one. The transition happens for $Ca\approx 0.09$.\label{fig:crosssections:sym}}
\end{figure}
